Predictions

### 2021-04-02 analysis/04_predictions.Rmd
Run all models and predictions!

---
title: "Predictions"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

@@

```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(targets)
library(here)

@@

  canary <- tar_read(panel_done_trimmed)
})
```

TODO:

- Split into training/test data
- Generate predictions
- Compare correctness, AUC, etc. for models with and without CS restrictions

>>>>

---
title: "Predictions"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

@@

```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(targets)
library(yardstick)
library(broom)
library(patchwork)
library(scales)
library(here)

@@

  canary_testing_lagged <- tar_read(panel_testing_lagged)

  # Load big list of models
  model_df <- tar_read(model_df) %>% 
    filter(str_detect(model, "baseline") | str_detect(model, "v2csreprss"))

  # Load actual model objects
  tar_load(c(m_pts_baseline_train, m_pts_v2csreprss_train,
             m_pts_total_train, m_pts_advocacy_train, 
             m_pts_entry_train, m_pts_funding_train,
             m_pts_baseline_rewb_train, m_pts_v2csreprss_rewb_train, 
             m_clphy_baseline_train, m_clphy_v2csreprss_train, 
             m_clphy_total_train, m_clphy_advocacy_train, 
             m_clphy_entry_train, m_clphy_funding_train,
             m_clpriv_baseline_train, m_clpriv_v2csreprss_train,
             m_clpriv_total_train, m_clpriv_advocacy_train, 
             m_clpriv_entry_train, m_clpriv_funding_train))
})

pts_levels <- levels(canary_testing_lagged$PTS_factor)

# Returns a data frame of predicted probabilities with actual = 1 
# when the response outcome happens
match_actual <- function(x, pred, actual) {
  pred %>% 
    select(fitted = {{x}}) %>% 
    mutate(actual = as.numeric(actual == colnames(pred[x])),
           plot_level = colnames(pred[x]))
}
```

```{r kludgy-bandaid-for-weird-data-issues}
# OOOOOOOF OH NO for some reason PTS_lead1 and all leaded V-Dem values missing
# for 2013 (i.e. 2014 values) and I don't want to go back and rerun all these
# models yet, so we do a temporary fix here and manually add the values from
# pts_clean and vdem_clean
withr::with_dir(here::here(), {
  tar_load(pts_clean)
  tar_load(vdem_clean)
})

pts_2014 <- pts_clean %>% 
  group_by(gwcode) %>% 
  mutate(across(c(PTS, PTS_factor),
                  list(lead1_2014 = ~lead(., 1)))) %>% 
  ungroup() %>% 
  filter(year == 2013) %>% 
  select(-PTS, -PTS_factor)

vdem_2014 <- vdem_clean %>% 
  group_by(gwcode) %>% 
  mutate(across(c(v2x_clphy, v2x_clpriv),
                  list(lead1_2014 = ~lead(., 1)))) %>% 
  ungroup() %>% 
  filter(year == 2013) %>% 
  select(gwcode, year, v2x_clphy_lead1_2014, v2x_clpriv_lead1_2014)

canary_testing_lagged <- canary_testing_lagged %>% 
  left_join(pts_2014, by = c("year", "gwcode")) %>% 
  left_join(vdem_2014, by = c("year", "gwcode")) %>% 
  mutate(PTS_lead1 = coalesce(PTS_lead1, PTS_lead1_2014),
         PTS_factor_lead1 = coalesce(PTS_factor_lead1, PTS_factor_lead1_2014),
         v2x_clphy_lead1 = coalesce(v2x_clphy_lead1, v2x_clphy_lead1_2014),
         v2x_clpriv_lead1 = coalesce(v2x_clpriv_lead1, v2x_clpriv_lead1_2014)) %>% 
  select(-contains("_2014"))
```

# E~1a~: NGO laws and political terror

```{r e1a-predictions, warning=FALSE}
seed_pts_prediction <- 5494  # From random.org

set.seed(seed_pts_prediction)
pred_pts_baseline <- predict(
  m_pts_baseline_train, 
  newdata = select(canary_testing_lagged, -PTS_factor_lead1),
  allow_new_levels = TRUE
) %>% 
  as_tibble() %>% 
  magrittr::set_colnames(pts_levels)

set.seed(seed_pts_prediction)
pred_pts_total <- predict(
  m_pts_total_train, 
  newdata = select(canary_testing_lagged, -PTS_factor_lead1),
  allow_new_levels = TRUE
) %>% 
  as_tibble() %>% 
  magrittr::set_colnames(pts_levels)

set.seed(seed_pts_prediction)
pred_pts_advocacy <- predict(
  m_pts_advocacy_train, 
  newdata = select(canary_testing_lagged, -PTS_factor_lead1),
  allow_new_levels = TRUE
) %>% 
  as_tibble() %>% 
  magrittr::set_colnames(pts_levels)

set.seed(seed_pts_prediction)
pred_pts_entry <- predict(
  m_pts_entry_train, 
  newdata = select(canary_testing_lagged, -PTS_factor_lead1),
  allow_new_levels = TRUE
) %>% 
  as_tibble() %>% 
  magrittr::set_colnames(pts_levels)

set.seed(seed_pts_prediction)
pred_pts_funding <- predict(
  m_pts_funding_train, 
  newdata = select(canary_testing_lagged, -PTS_factor_lead1),
  allow_new_levels = TRUE
) %>% 
  as_tibble() %>% 
  magrittr::set_colnames(pts_levels)
```

## Separation plots

### Baseline models

```{r e1a-separation-baseline}
# Match each column in pred_mat (predicted probabilities for each response level)
pred_long_list <- lapply(1:ncol(pred_pts_baseline), 
                         FUN = match_actual, 
                         pred = pred_pts_baseline, 
                         actual = as.character(canary_testing_lagged$PTS_factor_lead1))

plot_data <- pred_long_list %>% 
  bind_rows() %>% 
  filter(!is.nan(fitted), !is.na(actual)) %>% 
  group_by(plot_level) %>% 
  arrange(plot_level, fitted) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

markers <- plot_data %>% 
  group_by(plot_level) %>%
  summarize(obs = n(), expected = sum(fitted),
            marker = obs - expected + 1)

n_in_group <- nrow(pred_pts_baseline)

plot_data_condensed <- plot_data %>% 
  # rle(), dplyr-style: https://stackoverflow.com/a/33510765/120898
  mutate(rleid = (actual != lag(actual, 1, default = 0))) %>%
  mutate(rleid = cumsum(rleid)) %>% 
  group_by(plot_level, rleid) %>% 
  slice(1) %>% 
  group_by(plot_level) %>% 
  mutate(xmin = index, xmax = lead(index, default = n_in_group)) %>% 
  ungroup()

ggplot(plot_data_condensed) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 1, fill = as.factor(actual))) + 
  geom_line(aes(x = index, y = fitted), size = 0.3) +
  geom_point(data = markers, aes(x = marker, y = -0.05), shape = 17, size = 3) + 
  scale_fill_manual(values = c("grey90", "#2ECC40")) +
  guides(fill = FALSE) +
  facet_wrap(vars(plot_level), ncol = 1, scales = "free_x") +
  theme_void(base_family = "Inter Bold")
```

### Models with NGO laws

#### Total legal barriers

```{r e1a-separation-total}
# Match each column in pred_mat (predicted probabilities for each response level)
pred_long_list <- lapply(1:ncol(pred_pts_total), 
                         FUN = match_actual, 
                         pred = pred_pts_total, 
                         actual = as.character(canary_testing_lagged$PTS_factor_lead1))

plot_data <- pred_long_list %>% 
  bind_rows() %>% 
  filter(!is.nan(fitted), !is.na(actual)) %>% 
  group_by(plot_level) %>% 
  arrange(plot_level, fitted) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

markers <- plot_data %>% 
  group_by(plot_level) %>%
  summarize(obs = n(), expected = sum(fitted),
            marker = obs - expected + 1)

n_in_group <- nrow(pred_pts_total)

plot_data_condensed <- plot_data %>% 
  # rle(), dplyr-style: https://stackoverflow.com/a/33510765/120898
  mutate(rleid = (actual != lag(actual, 1, default = 0))) %>%
  mutate(rleid = cumsum(rleid)) %>% 
  group_by(plot_level, rleid) %>% 
  slice(1) %>% 
  group_by(plot_level) %>% 
  mutate(xmin = index, xmax = lead(index, default = n_in_group)) %>% 
  ungroup()

ggplot(plot_data_condensed) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 1, fill = as.factor(actual))) + 
  geom_line(aes(x = index, y = fitted), size = 0.3) +
  geom_point(data = markers, aes(x = marker, y = -0.05), shape = 17, size = 3) + 
  scale_fill_manual(values = c("grey90", "#2ECC40")) +
  guides(fill = FALSE) +
  facet_wrap(vars(plot_level), ncol = 1, scales = "free_x") +
  theme_void(base_family = "Inter Bold")
```

#### Barriers to advocacy

```{r e1a-separation-advocacy}
# Match each column in pred_mat (predicted probabilities for each response level)
pred_long_list <- lapply(1:ncol(pred_pts_advocacy), 
                         FUN = match_actual, 
                         pred = pred_pts_advocacy, 
                         actual = as.character(canary_testing_lagged$PTS_factor_lead1))

plot_data <- pred_long_list %>% 
  bind_rows() %>% 
  filter(!is.nan(fitted), !is.na(actual)) %>% 
  group_by(plot_level) %>% 
  arrange(plot_level, fitted) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

markers <- plot_data %>% 
  group_by(plot_level) %>%
  summarize(obs = n(), expected = sum(fitted),
            marker = obs - expected + 1)

n_in_group <- nrow(pred_pts_advocacy)

plot_data_condensed <- plot_data %>% 
  # rle(), dplyr-style: https://stackoverflow.com/a/33510765/120898
  mutate(rleid = (actual != lag(actual, 1, default = 0))) %>%
  mutate(rleid = cumsum(rleid)) %>% 
  group_by(plot_level, rleid) %>% 
  slice(1) %>% 
  group_by(plot_level) %>% 
  mutate(xmin = index, xmax = lead(index, default = n_in_group)) %>% 
  ungroup()

ggplot(plot_data_condensed) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 1, fill = as.factor(actual))) + 
  geom_line(aes(x = index, y = fitted), size = 0.3) +
  geom_point(data = markers, aes(x = marker, y = -0.05), shape = 17, size = 3) + 
  scale_fill_manual(values = c("grey90", "#2ECC40")) +
  guides(fill = FALSE) +
  facet_wrap(vars(plot_level), ncol = 1, scales = "free_x") +
  theme_void(base_family = "Inter Bold")
```

#### Barriers to entry

```{r e1a-separation-entry}
# Match each column in pred_mat (predicted probabilities for each response level)
pred_long_list <- lapply(1:ncol(pred_pts_entry), 
                         FUN = match_actual, 
                         pred = pred_pts_entry, 
                         actual = as.character(canary_testing_lagged$PTS_factor_lead1))

plot_data <- pred_long_list %>% 
  bind_rows() %>% 
  filter(!is.nan(fitted), !is.na(actual)) %>% 
  group_by(plot_level) %>% 
  arrange(plot_level, fitted) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

markers <- plot_data %>% 
  group_by(plot_level) %>%
  summarize(obs = n(), expected = sum(fitted),
            marker = obs - expected + 1)

n_in_group <- nrow(pred_pts_entry)

plot_data_condensed <- plot_data %>% 
  # rle(), dplyr-style: https://stackoverflow.com/a/33510765/120898
  mutate(rleid = (actual != lag(actual, 1, default = 0))) %>%
  mutate(rleid = cumsum(rleid)) %>% 
  group_by(plot_level, rleid) %>% 
  slice(1) %>% 
  group_by(plot_level) %>% 
  mutate(xmin = index, xmax = lead(index, default = n_in_group)) %>% 
  ungroup()

ggplot(plot_data_condensed) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 1, fill = as.factor(actual))) + 
  geom_line(aes(x = index, y = fitted), size = 0.3) +
  geom_point(data = markers, aes(x = marker, y = -0.05), shape = 17, size = 3) + 
  scale_fill_manual(values = c("grey90", "#2ECC40")) +
  guides(fill = FALSE) +
  facet_wrap(vars(plot_level), ncol = 1, scales = "free_x") +
  theme_void(base_family = "Inter Bold")
```

#### Barriers to funding

```{r e1a-separation-funding}
# Match each column in pred_mat (predicted probabilities for each response level)
pred_long_list <- lapply(1:ncol(pred_pts_funding), 
                         FUN = match_actual, 
                         pred = pred_pts_funding, 
                         actual = as.character(canary_testing_lagged$PTS_factor_lead1))

plot_data <- pred_long_list %>% 
  bind_rows() %>% 
  filter(!is.nan(fitted), !is.na(actual)) %>% 
  group_by(plot_level) %>% 
  arrange(plot_level, fitted) %>% 
  mutate(index = row_number()) %>% 
  ungroup()

markers <- plot_data %>% 
  group_by(plot_level) %>%
  summarize(obs = n(), expected = sum(fitted),
            marker = obs - expected + 1)

n_in_group <- nrow(pred_pts_funding)

plot_data_condensed <- plot_data %>% 
  # rle(), dplyr-style: https://stackoverflow.com/a/33510765/120898
  mutate(rleid = (actual != lag(actual, 1, default = 0))) %>%
  mutate(rleid = cumsum(rleid)) %>% 
  group_by(plot_level, rleid) %>% 
  slice(1) %>% 
  group_by(plot_level) %>% 
  mutate(xmin = index, xmax = lead(index, default = n_in_group)) %>% 
  ungroup()

ggplot(plot_data_condensed) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = 1, fill = as.factor(actual))) + 
  geom_line(aes(x = index, y = fitted), size = 0.3) +
  geom_point(data = markers, aes(x = marker, y = -0.05), shape = 17, size = 3) + 
  scale_fill_manual(values = c("grey90", "#2ECC40")) +
  guides(fill = FALSE) +
  facet_wrap(vars(plot_level), ncol = 1, scales = "free_x") +
  theme_void(base_family = "Inter Bold")
```

## Prediction diagnostics

```{r e1a-improved-cases}
pred_pts_baseline_categories <- pred_pts_baseline %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  group_by(rowname) %>% 
  filter(rank(-value, ties.method = "last") == 1) %>% 
  ungroup() %>% 
  select(predicted = name)

pred_pts_total_categories <- pred_pts_total %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  group_by(rowname) %>% 
  filter(rank(-value, ties.method = "last") == 1) %>% 
  ungroup() %>% 
  select(predicted = name)

pred_pts_advocacy_categories <- pred_pts_advocacy %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  group_by(rowname) %>% 
  filter(rank(-value, ties.method = "last") == 1) %>% 
  ungroup() %>% 
  select(predicted = name)

pred_pts_entry_categories <- pred_pts_entry %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  group_by(rowname) %>% 
  filter(rank(-value, ties.method = "last") == 1) %>% 
  ungroup() %>% 
  select(predicted = name)

pred_pts_funding_categories <- pred_pts_funding %>% 
  rownames_to_column() %>% 
  pivot_longer(-rowname) %>% 
  group_by(rowname) %>% 
  filter(rank(-value, ties.method = "last") == 1) %>% 
  ungroup() %>% 
  select(predicted = name)

pts_cat_preds <- canary_testing_lagged %>% 
  select(country, year, actual = PTS_factor_lead1) %>% 
  mutate(pred_baseline = pred_pts_baseline_categories$predicted,
         pred_total = pred_pts_total_categories$predicted,
         pred_advocacy = pred_pts_advocacy_categories$predicted,
         pred_entry = pred_pts_entry_categories$predicted,
         pred_funding = pred_pts_funding_categories$predicted) %>% 
  filter(!is.na(actual)) %>%
  mutate(is_correct_baseline = actual == pred_baseline,
         is_correct_total = actual == pred_total,
         is_correct_advocacy = actual == pred_advocacy,
         is_correct_entry = actual == pred_entry,
         is_correct_funding = actual == pred_funding)

# Improved predictions
pts_improved_total <- pts_cat_preds %>% 
  filter(!is_correct_baseline & is_correct_total) %>% 
  select(Country = country, Year = year, Actual = actual,
         `CS repression included` = pred_total,
         `Baseline` = pred_baseline)

pts_improved_advocacy <- pts_cat_preds %>% 
  filter(!is_correct_baseline & is_correct_advocacy) %>% 
  select(Country = country, Year = year, Actual = actual,
         `CS repression included` = pred_advocacy,
         `Baseline` = pred_baseline)

pts_improved_entry <- pts_cat_preds %>% 
  filter(!is_correct_baseline & is_correct_entry) %>% 
  select(Country = country, Year = year, Actual = actual,
         `CS repression included` = pred_entry,
         `Baseline` = pred_baseline)

pts_improved_funding <- pts_cat_preds %>% 
  filter(!is_correct_baseline & is_correct_funding) %>% 
  select(Country = country, Year = year, Actual = actual,
         `CS repression included` = pred_funding,
         `Baseline` = pred_baseline)

pts_improved_total %>% 
  kbl(caption = "Cases improved with total legal barriers") %>% 
  kable_styling()

pts_improved_advocacy %>% 
  kbl(caption = "Cases improved with barriers to advocacy") %>% 
  kable_styling()

pts_improved_entry %>% 
  kbl(caption = "Cases improved with barriers to entry") %>% 
  kable_styling()

pts_improved_funding %>% 
  kbl(caption = "Cases improved with barriers to funding") %>% 
  kable_styling()
```

```{r e1a-correct-not-correct}
correct_lookup <- tribble(
  ~name, ~nice_name,
  "is_correct_total", "Total legal barriers included",
  "is_correct_advocacy", "Barriers to advocacy included",
  "is_correct_entry", "Barriers to entry included",
  "is_correct_funding", "Barriers to funding included",
  "is_correct_baseline", "Baseline"
) %>% mutate(nice_name = fct_inorder(nice_name, ordered = TRUE))

pts_pred_table <- pts_cat_preds %>% 
  pivot_longer(starts_with("is_correct")) %>% 
  group_by(name, value) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  left_join(correct_lookup, by = "name") %>% 
  mutate(Prediction = ifelse(value == TRUE, "Correct", "Wrong")) %>% 
  arrange(nice_name) %>% 
  select(-name, -value) %>% 
  pivot_wider(names_from = "nice_name", values_from = "n")

pts_pred_table %>% 
  kbl(align = "lcccc") %>% 
  kable_styling()
```

\

# E~1b~: NGO laws and physical violence

```{r e1b-predictions, warning=FALSE}
seed_clphy_prediction <- 7285  # From random.org

set.seed(seed_clphy_prediction)
pred_clphy_baseline <- predict(m_clphy_baseline_train, 
                               newdata = select(canary_testing_lagged, -v2x_clphy_lead1),
                               allow_new_levels = TRUE)

set.seed(seed_clphy_prediction)
pred_clphy_total <- predict(m_clphy_total_train, 
                            newdata = select(canary_testing_lagged, -v2x_clphy_lead1),
                            allow_new_levels = TRUE)

set.seed(seed_clphy_prediction)
pred_clphy_advocacy <- predict(m_clphy_advocacy_train, 
                               newdata = select(canary_testing_lagged, -v2x_clphy_lead1),
                               allow_new_levels = TRUE)

set.seed(seed_clphy_prediction)
pred_clphy_entry <- predict(m_clphy_entry_train, 
                            newdata = select(canary_testing_lagged, -v2x_clphy_lead1),
                            allow_new_levels = TRUE)

set.seed(seed_clphy_prediction)
pred_clphy_funding <- predict(m_clphy_funding_train, 
                              newdata = select(canary_testing_lagged, -v2x_clphy_lead1),
                              allow_new_levels = TRUE)

pred_clphy_baseline_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clphy_lead1) %>% 
  bind_cols(as_tibble(pred_clphy_baseline)) %>% 
  filter(!is.na(v2x_clphy_lead1))

pred_clphy_total_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clphy_lead1) %>% 
  bind_cols(as_tibble(pred_clphy_total)) %>% 
  filter(!is.na(v2x_clphy_lead1))

pred_clphy_advocacy_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clphy_lead1) %>% 
  bind_cols(as_tibble(pred_clphy_advocacy)) %>% 
  filter(!is.na(v2x_clphy_lead1))

pred_clphy_entry_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clphy_lead1) %>% 
  bind_cols(as_tibble(pred_clphy_entry)) %>% 
  filter(!is.na(v2x_clphy_lead1))

pred_clphy_funding_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clphy_lead1) %>% 
  bind_cols(as_tibble(pred_clphy_funding)) %>% 
  filter(!is.na(v2x_clphy_lead1))
```

## Improvement from baseline models

```{r e1b-pred-plots, warning=FALSE}
plot_baseline <- ggplot(pred_clphy_baseline_test, 
                        aes(x = v2x_clphy_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Baseline",
       x = "Actual physical violence index (t + 1)",
       y = "Predicted physical violence index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_total <- ggplot(pred_clphy_total_test, 
                     aes(x = v2x_clphy_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Total legal barriers",
       x = "Actual physical violence index (t + 1)",
       y = "Predicted physical violence index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_advocacy <- ggplot(pred_clphy_advocacy_test, 
                        aes(x = v2x_clphy_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to advocacy",
       x = "Actual physical violence index (t + 1)",
       y = "Predicted physical violence index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_entry <- ggplot(pred_clphy_entry_test, 
                     aes(x = v2x_clphy_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to entry",
       x = "Actual physical violence index (t + 1)",
       y = "Predicted physical violence index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_funding <- ggplot(pred_clphy_funding_test, 
                       aes(x = v2x_clphy_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to funding",
       x = "Actual physical violence index (t + 1)",
       y = "Predicted physical violence index (t + 1)") +
  coord_equal() +
  theme_ngo()

(plot_baseline | plot_total) / (plot_advocacy | plot_entry | plot_funding)
```

```{r e1b-metrics}
lotsa_metrics <- metric_set(rmse, rsq, mae, ccc)

bind_rows(lotsa_metrics(pred_clphy_baseline_test, 
                        truth = v2x_clphy_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Baseline"),
          lotsa_metrics(pred_clphy_total_test, 
                        truth = v2x_clphy_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Total legal barriers"),
          lotsa_metrics(pred_clphy_advocacy_test, 
                        truth = v2x_clphy_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Barriers to advocacy"),
          lotsa_metrics(pred_clphy_entry_test, 
                        truth = v2x_clphy_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Barriers to entry"),
          lotsa_metrics(pred_clphy_funding_test, 
                        truth = v2x_clphy_lead1, 
                        estimate = Estimate) %>% 
            mutate(Model = "Barriers to funding")) %>% 
  select(Model, .metric, .estimate) %>% 
  pivot_wider(names_from = ".metric", values_from = ".estimate")
```

## Most improved countries

### Total legal barriers

```{r e1b-most-improved-total}
most_improved_clphy_total <- bind_cols(
  pred_clphy_baseline_test %>% 
    mutate(error_baseline = Estimate - v2x_clphy_lead1) %>% 
    select(country, gwcode, year, v2x_clphy_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_clphy_total_test %>% 
    mutate(error_total = Estimate - v2x_clphy_lead1) %>% 
    select(estimate_total = Estimate, error_total)
) %>% 
  mutate(improved = error_total < 0 & abs(error_total) < abs(error_baseline)) %>% 
  mutate(diff = error_total - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_clphy_total)

most_improved_clphy_total %>% 
  select(Country = country, Year = year, 
         `Actual physical violence index (t + 1)` =  v2x_clphy_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_total,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

### Barriers to advocacy

```{r e1b-most-improved-advocacy}
most_improved_clphy_advocacy <- bind_cols(
  pred_clphy_baseline_test %>% 
    mutate(error_baseline = Estimate - v2x_clphy_lead1) %>% 
    select(country, gwcode, year, v2x_clphy_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_clphy_advocacy_test %>% 
    mutate(error_advocacy = Estimate - v2x_clphy_lead1) %>% 
    select(estimate_advocacy = Estimate, error_advocacy)
) %>% 
  mutate(improved = error_advocacy < 0 & abs(error_advocacy) < abs(error_baseline)) %>% 
  mutate(diff = error_advocacy - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_clphy_advocacy)

most_improved_clphy_advocacy %>% 
  select(Country = country, Year = year, 
         `Actual physical violence index (t + 1)` =  v2x_clphy_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_advocacy,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

### Barriers to entry

```{r e1b-most-improved-entry}
most_improved_clphy_entry <- bind_cols(
  pred_clphy_baseline_test %>% 
    mutate(error_baseline = Estimate - v2x_clphy_lead1) %>% 
    select(country, gwcode, year, v2x_clphy_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_clphy_entry_test %>% 
    mutate(error_entry = Estimate - v2x_clphy_lead1) %>% 
    select(estimate_entry = Estimate, error_entry)
) %>% 
  mutate(improved = error_entry < 0 & abs(error_entry) < abs(error_baseline)) %>% 
  mutate(diff = error_entry - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_clphy_entry)

most_improved_clphy_entry %>% 
  select(Country = country, Year = year, 
         `Actual physical violence index (t + 1)` =  v2x_clphy_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_entry,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

### Barriers to funding

```{r e1b-most-improved-funding}
most_improved_clphy_funding <- bind_cols(
  pred_clphy_baseline_test %>% 
    mutate(error_baseline = Estimate - v2x_clphy_lead1) %>% 
    select(country, gwcode, year, v2x_clphy_lead1, 
           estimate_baseline = Estimate, error_baseline),
  pred_clphy_funding_test %>% 
    mutate(error_funding = Estimate - v2x_clphy_lead1) %>% 
    select(estimate_funding = Estimate, error_funding)
) %>% 
  mutate(improved = error_funding < 0 & abs(error_funding) < abs(error_baseline)) %>% 
  mutate(diff = error_funding - error_baseline) %>% 
  mutate(abs_diff = abs(diff)) %>% 
  filter(improved) %>% 
  arrange(desc(abs_diff))

nrow(most_improved_clphy_funding)

most_improved_clphy_funding %>% 
  select(Country = country, Year = year, 
         `Actual physical violence index (t + 1)` =  v2x_clphy_lead1,
         `Predicted (baseline)` = estimate_baseline,
         `Predicted (CS repression)` = estimate_funding,
         `|CS repression − baseline|` = abs_diff) %>% 
  head(5) %>% 
  kbl(digits = 3, align = "lccccc") %>% 
  kable_styling()
```

\

# E~1c~: NGO laws and civil liberties

```{r e1c-predictions, warning=FALSE}
seed_clpriv_prediction <- 6670  # From random.org

set.seed(seed_clpriv_prediction)
pred_clpriv_baseline <- predict(m_clpriv_baseline_train, 
                                newdata = select(canary_testing_lagged, -v2x_clpriv_lead1),
                                allow_new_levels = TRUE)

set.seed(seed_clpriv_prediction)
pred_clpriv_total <- predict(m_clpriv_total_train, 
                             newdata = select(canary_testing_lagged, -v2x_clpriv_lead1),
                             allow_new_levels = TRUE)

set.seed(seed_clpriv_prediction)
pred_clpriv_advocacy <- predict(m_clpriv_advocacy_train, 
                                newdata = select(canary_testing_lagged, -v2x_clpriv_lead1),
                                allow_new_levels = TRUE)

set.seed(seed_clpriv_prediction)
pred_clpriv_entry <- predict(m_clpriv_entry_train, 
                             newdata = select(canary_testing_lagged, -v2x_clpriv_lead1),
                             allow_new_levels = TRUE)

set.seed(seed_clpriv_prediction)
pred_clpriv_funding <- predict(m_clpriv_funding_train, 
                               newdata = select(canary_testing_lagged, -v2x_clpriv_lead1),
                               allow_new_levels = TRUE)

pred_clpriv_baseline_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clpriv_lead1) %>% 
  bind_cols(as_tibble(pred_clpriv_baseline)) %>% 
  filter(!is.na(v2x_clpriv_lead1))

pred_clpriv_total_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clpriv_lead1) %>% 
  bind_cols(as_tibble(pred_clpriv_total)) %>% 
  filter(!is.na(v2x_clpriv_lead1))

pred_clpriv_advocacy_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clpriv_lead1) %>% 
  bind_cols(as_tibble(pred_clpriv_advocacy)) %>% 
  filter(!is.na(v2x_clpriv_lead1))

pred_clpriv_entry_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clpriv_lead1) %>% 
  bind_cols(as_tibble(pred_clpriv_entry)) %>% 
  filter(!is.na(v2x_clpriv_lead1))

pred_clpriv_funding_test <- canary_testing_lagged %>% 
  select(country, gwcode, year, v2x_clpriv_lead1) %>% 
  bind_cols(as_tibble(pred_clpriv_funding)) %>% 
  filter(!is.na(v2x_clpriv_lead1))
```

## Improvement from baseline models

```{r e1c-pred-plots, warning=FALSE}
plot_baseline <- ggplot(pred_clpriv_baseline_test, 
                        aes(x = v2x_clpriv_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Baseline",
       x = "Actual private civil liberties index (t + 1)",
       y = "Predicted private civil liberties index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_total <- ggplot(pred_clpriv_total_test, 
                     aes(x = v2x_clpriv_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Total legal barriers",
       x = "Actual private civil liberties index (t + 1)",
       y = "Predicted private civil liberties index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_advocacy <- ggplot(pred_clpriv_advocacy_test, 
                        aes(x = v2x_clpriv_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to advocacy",
       x = "Actual private civil liberties index (t + 1)",
       y = "Predicted private civil liberties index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_entry <- ggplot(pred_clpriv_entry_test, 
                     aes(x = v2x_clpriv_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to entry",
       x = "Actual private civil liberties index (t + 1)",
       y = "Predicted private civil liberties index (t + 1)") +
  coord_equal() +
  theme_ngo()

plot_funding <- ggplot(pred_clpriv_funding_test, 
                       aes(x = v2x_clpriv_lead1, y = Estimate)) +
  geom_abline(lty = 2) +
  geom_point(alpha = 0.5) +
  labs(title = "Barriers to funding",
       x = "Actual private civil liberties index (t + 1)",
       y = "Predicted private civil liberties index (t + 1)") +
  coord_equal() +
  theme_ngo()

(plot_baseline | plot_total) / (plot_advocacy | plot_entry | plot_funding)
```

```{r e1c-metrics}
lotsa_metrics <- metric_set(rmse, rsq, mae, ccc)
